### 2013-06-20 cv-body.tex
Bankruptcy commit since moving to Linode

\section{Education}

\begin{description}
\item[2011--13]
Ph.D.\ Candidate, Simon Fraser University, Burnaby, Canada Supervisors:
Dr.\ Nicholas K. Dulvy and Dr.\ Andrew B. Cooper\\Canadian Fulbright
Scholar to the University of Washington in 2012--13
\item[2010]
M.Sc.\ Biology, Dalhousie University, Halifax, Canada\\Thesis title:
Trends, drivers, and ecosystem effects of expanding global invertebrate
fisheries. Supervisor: Dr.\ Heike K. Lotze.
(\href{papers/Anderson_MSc_2010.pdf}{PDF})\\Canadian
\href{http://goo.gl/nA1zE}{Governor General's Academic Gold Medal} for
the top-ranked Master's Natural Sciences and Engineering thesis at
Dalhousie University in 2010
\item[2004--07]
B.Sc.\ (Hons.) Environmental Science, Dalhousie University, Halifax,
Canada\\Thesis title: How sustainable are emerging low-trophic level
fisheries on the Scotian Shelf? Supervisors: Dr.\ Heike K. Lotze and
Dr.\ Ransom A. Myers. (\href{papers/Anderson_HSc_2006.pdf}{PDF})\\First
Class Honours \& Environmental Programmes Honour Society Medal
\item[2001--03]
B.Sc.\ Candidate (Hons.) Kinesiology, University of Waterloo, Waterloo,
Canada Transferred to Dalhousie University in 2004.
\end{description}

\section{Publications}

\begin{description}
\item[2012]
Phillis, C.C.*, S.M. O'Regan*, S.J. Green*, J.E.B. Bruce*, \textbf{S.C.
Anderson}, J.N. Linton, Earth2Ocean Research Derby, B. Favaro. Multiple
pathways to conservation success. Conservation Letters. Early view
online. \url{http://dx.doi.org/10.1111/j.1755-263X.2012.00294.x}.
(*\emph{Authors contributed equally, listed in reverse alphabetical
order})
\item[2012]
Harnik, P.G., H.K. Lotze, \textbf{S.C. Anderson}, Z.V. Finkel, S.
Finnegan, D.R. Lindberg, L.H. Liow, R. Lockwood, C.R. McClain, J.L.
McGuire, A. O'Dea, J.M. Pandolfi, C. Simpson, D.P. Tittensor.
Extinctions in ancient and modern seas. Trends in Ecology and Evolution.
27(11): 608--617. \url{http://dx.doi.org/10.1016/j.tree.2012.07.010}.
\item[2012]
\textbf{Anderson, S.C.}, T.A. Branch, D. Ricard, H.K. Lotze. Assessing
global marine fishery status with a revised dynamic catch-based method
and stock-assessment reference points. ICES Journal of Marine Science.
69(8): 1491--1500. \url{http://dx.doi.org/10.1093/icesjms/fss105}.
(\href{http://icesjms.oxfordjournals.org/cgi/reprint/fss105?ijkey=pFRzMGEDycomI3X\&keytype=ref}{PDF})
\item[2011]
\textbf{Anderson, S.C.}, B.G. Farmer, F. Ferretti, A.L.S. Houde, and
J.A. Hutchings. Correlates of vertebrate extinction risk in Canada.
BioScience. 61(7): 538--549.
\url{http://dx.doi.org/10.1525/bio.2011.61.7.8}.
(\href{papers/Anderson_etal_2011_BioScience_with_supplement.pdf}{PDF})
\item[2011]
\textbf{Anderson, S.C.}, J. Mills Flemming, R. Watson, H.K. Lotze.
Serial exploitation of global sea cucumber fisheries. Fish and
Fisheries. 12(3): 317-339.
\url{http://dx.doi.org/10.1111/j.1467-2979.2010.00397.x}.
(\href{papers/Anderson_etal_2011_seacucumbers_with_supplement.pdf}{PDF})
\href{http://www.sciencemag.org/content/331/6014/129.1.full}{Featured in
Science} on January 14 2011.
\item[2011]
Boudreau, S.A., \textbf{S.C. Anderson}, B. Worm. Top-down interactions
and temperature control of snow crab abundance in the northwest
Atlantic. Marine Ecology Progress Series. 429: 169--183.
\url{http://dx.doi.org/10.3354/meps09081}.
\item[2011]
\textbf{Anderson, S.C.}, J. Mills Flemming, R. Watson, H.K. Lotze. Rapid
global expansion of invertebrate fisheries: trends, drivers, and
ecosystem effects. PLoS ONE. 6(3): e14735.
\url{http://dx.doi.org/10.1371/journal.pone.0014735}
\item[2008]
\textbf{Anderson, S.C.}, H.K. Lotze, and N.L. Shackell. Evaluating the
knowledge base for expanding low-trophic-level fisheries in Atlantic
Canada. Can. J. Fish. Aquat. Sci. 65(12): 2553--2571.
\url{http://dx.doi.org/10.1139/F08-156}.
(\href{papers/Anderson_etal_2008_knowledge.pdf}{PDF}) Eighth most-read
CJFAS article in 2008.
\end{description}

\section{Submitted manuscripts}

\begin{description}
\item[2013]
\textbf{Anderson, S.C.}, A.B. Cooper, N.K. Dulvy. Ecological prophets:
Quantifying metapopulation portfolio effects. Submitted with revisions
to Methods in Ecology and Evolution.
\item[2013]
O'Regan, S.M., W.J. Palen, \textbf{S.C. Anderson}. Climate warming
mediates negative impacts of rapid pond drying for three amphibian
species. Submitted to Ecology.
\end{description}

\section{Awards and scholarships}

\begin{description}
\item[2014]
Graduate Fellowship, Simon Fraser University
\item[2012--13]
\href{http://www.fulbright.ca/programs/canadian-students/2012-2013-canadian-fulbright-students.html}{Canadian
Fulbright Scholar award} to the University of Washington in 2012--13
\item[2011]
Canadian \href{http://goo.gl/nA1zE}{Governor General's Academic Gold
Medal} for the top-ranked Master's Natural Sciences and Engineering
thesis at Dalhousie University in 2010
\item[2011--14]
Provost Prize of Distinction, Simon Fraser University
\item[2011--14]
Natural Sciences and Engineering Research Council of Canada Postgraduate
Scholarship (Doctoral)
\item[2007--10]
Faculty Research Grant Scholarship, Dalhousie University
\item[2007--09]
Graduate Studies Scholarship, Dalhousie University
\item[2007]
Environmental Programmes Honour Society Medal, Dalhousie University
\item[2001--03]
Academic All-Canadian, University of Waterloo
\item[2001--02]
Dean's Entrance Scholarship, University of Waterloo
\end{description}

\section{Invited talks and conference presentations}

\begin{description}
\item[2013]
\textbf{Anderson, S.C.}, A.B. Cooper, N.K. Dulvy. False prophets: The
challenges of quantifying ecological portfolios
(\href{talks/2013/PE_SAFS_quantsem.pdf}{slides}), Quantitative Seminar,
School of Aquatic and Fishery Sciences, University of Washington,
Seattle, WA, United States.
\item[2012]
\textbf{Anderson, S.C.}, A.B. Cooper, N.K. Dulvy. False prophets: The
ecological portfolio effect overestimates the benefit of diversity.
Branch Lab, School of Aquatic and Fishery Sciences, University of
Washington, Seattle, WA, United States.
\item[2012]
\textbf{Anderson, S.C.}, A.B. Cooper, N.K. Dulvy. Metapopulation
dynamics and the generality of the ecological portfolio effect. Les
Ecologistes Departmental Seminar, Simon Fraser University, Burnaby, BC,
Canada.
\item[2010]
\textbf{Anderson, S.C.} The rise of invertebrates, the fall of sea
cucumbers, and the risk of maturing late. Earth2Ocean Research Group,
Simon Fraser University, Burnaby, BC, Canada.
\item[2009]
\textbf{Anderson, S.C.}, J. Mills Flemming, R. Watson, H.K. Lotze.
Ecosystem impacts of the global expansion of invertebrate fisheries.
International Oceans Past II Conference \emph{Multidisciplinary
Perspectives on the History and Future of Marine Animal Populations}.
Vancouver, BC, Canada.
\item[2009]
\textbf{Anderson, S.C.}, J. Mills Flemming, R. Watson, H.K. Lotze.
Global invertebrate fisheries: trends and consequences. NCEAS (National
Center for Ecological Analysis and Synthesis) Working Group
\emph{Finding Common Ground in Marine Conservation and Management}.
Santa Barbara, CA, USA.
\item[2009]
\textbf{Anderson, S.C.}, H.K. Lotze, N.L. Shackell. Evaluating the
knowledge base for expanding low-trophic level fisheries in Atlantic
Canada. Harvest Fisheries Seminar Series. Bedford Institute of
Oceanography, Fisheries and Oceans, Dartmouth, NS, Canada.
\item[2008]
\textbf{Anderson, S.C.}, H.K. Lotze, N.L. Shackell. Evaluating the
knowledge base for expanding low-trophic level fisheries in Atlantic
Canada. Departmental Seminar, Biology Department, Dalhousie University,
Canada.
\end{description}

\section{Contract employment}

\begin{description}
\item[2011--13]
Database management, data analysis, and visualization for the project
\emph{Global Shark Abundance Baselines} (Hopkins Marine Station,
Stanford University, CA, United States) with F. Micheli, F. Ferretti,
N.K. Dulvy, H.K. Lotze, and B. Worm.
\item[2010]
Assessment of Arctic surfclam on Banquereau Bank using industry
collected data for Clearwater Seafoods, Bedford, NS, Canada.
\item[2010]
Modelling length-frequency distribution changes of Atlantic halibut for
NAFO Divisions 4VWX5Z and 3NOP for Dr.\ K. Trzcinski, Population Ecology
Division, Bedford Institute of Oceanography, Fisheries and Oceans,
Dartmouth, NS, Canada.
\end{description}

\section{Working groups and workshops}

\begin{description}
\item[2011--13]
NESCent (National Evolutionary Synthesis Center, Durham, NC) Working
Group \emph{Determinants of Extinction in Ancient and Modern Seas} led
by Paul Harnik, Seth Finnegan, and Rowan Lockwood.
(\href{http://www.nescent.org/science/awards_summary.php?id=256}{URL})
\item[2010]
Atlantic Halibut Assessment Science Peer Review Meeting, Fisheries and
Oceans, Dartmouth, NS, Canada.
\item[2007--09]
NCEAS (National Center for Ecological Analysis and Synthesis, Santa
Barbara, CA) Distributed Graduate Seminar, in association with the
Working Group \emph{Finding Common Ground in Marine Conservation and
Management} led by Ray Hilborn and Boris Worm.
(\href{http://www.nceas.ucsb.edu/projects/12307}{URL})
\item[2008]
Workshop on Canadian Science and Management Strategies for Sea Cucumber
(\emph{Cucumaria frondosa}), Fisheries and Oceans, Dartmouth, NS, Canada
(work presented \emph{in absentia}).
\item[2007]
Workshop on Canadian Science and Management Strategies for Atlantic
Hagfish, Fisheries and Oceans, Dartmouth, NS, Canada.
(\href{http://www.dfo-mpo.gc.ca/CSAS/Csas/Publications/Pro-CR/2009/2009_009_e.htm}{URL})
\end{description}

\section{Selected courses and training}

\begin{description}
\item[Present]
Proficient in statistical analyses and data management using R, WinBUGS,
JAGS, Sweave, knitr, LaTeX, Perl, PostgreSQL, Git.
\item[2013]
Quantitative Modelling in Ecosystem Based Management, R. Hilborn, School
of Aquatic and Fishery Sciences, University of Washington (auditing).
\item[2013]
Applied Time Series Analysis; E. Holmes, M. Scheuerell, and E. Ward;
School of Aquatic and Fishery Sciences; University of Washington
(audited).
\item[2012]
Numerical Computing for the Natural Resources, Dr.\ A.E. Punt, School of
Aquatic and Fishery Sciences, University of Washington (audited).
\item[2012]
Risk Assessment and Decision Analysis for Management of Natural
Resources, Dr.\ R.M. Peterman, School of Resource and Environmental
Management, Simon Fraser University.
\item[2011]
Presenting Data and Information, Edward Tufte, Seattle, WA, United
States.
\item[2009]
Technical Expertise in Stock Assessment: Model Fitting and
Interpretation by Dr.\ C. Needle, chair of the International Council for
the Exploration of the Sea working group on methods of fish stock
assessment, Dalhousie University.
\item[2008]
Data Analysis, Dr.\ J. Mills Flemming, Department of Mathematics \&
Statistics, Dalhousie University (audited).
\end{description}

\section{Teaching}

\begin{description}
\item[2012]
Introduction to ggplot2.
(\href{courses/12-ggplot2/ggplot2_notes.pdf}{notes},
\href{courses/12-ggplot2/ggplot2_slides_with_examples.pdf}{slides})
Lecture for FISH 507H: Beautiful Graphics in R, School of Aquatic and
Fishery Sciences, University of Washington, Seattle, WA, United States
\item[2012]
Workshop on the R package plyr.
(\href{courses/12-plyr/plyr_2012.pdf}{notes},
\href{courses/12-plyr/plyr_2012_slides.pdf}{slides},
\href{courses/12-plyr/plyr_2012_examples.html}{examples}) Branch Lab,
School of Aquatic and Fishery Sciences, University of Washington,
Seattle, WA, United States
\item[2011]
Multipanel plotting in R with base graphics.
(\href{courses/11-multipanel/multipanel.pdf}{notes},
\href{courses/11-multipanel/multipanel-slides.pdf}{slides}). Earth2Ocean
Research Group, Simon Fraser University, Burnaby, BC, Canada.
\item[2011]
A brief introduction to R. (\href{courses/11-rintro/RIntro.pdf}{notes},
\href{courses/11-rintro/RIntro.R}{workshop code}). Earth2Ocean Research
Group, Simon Fraser University, Burnaby, BC, Canada.
\item[2011]
Workshop on the R package plyr (\href{courses/11-plyr/plyr.pdf}{notes},
\href{courses/11-plyr/plyr-slides.pdf}{slides}). Earth2Ocean Research
Group, Simon Fraser University, Burnaby, BC, Canada.
\item[2007--08]
Teaching Assistant, Organismal Biology and Ecology, Dalhousie
University, two semesters (BIOL 1021).
\item[2007--08]
Teaching Assistant, Marine Mammology, Dalhousie University (BIOL 4060).
\end{description}

\section{Reviews}

Reviewer for Ecology, Conservation Biology, Fish and Fisheries,
International Journal of Tropical Biology and Conservation, Journal of
Environmental Management, Aquatic Conservation