### 2013-10-20 _knitr-posts/2013-10-19-reshape.Rmd
Add reshape .Rmd file

---
layout: post
title: An Introduction to reshape2
subtitle: Reshaping data easily with the reshape2 R package.
published: true
---

[`reshape2`](http://cran.r-project.org/web/packages/reshape2/index.html) is an
`R` package written by [Hadley Wickham](http://had.co.nz) that makes it easy
to transform data between wide and long formats.

What makes data wide or long? 
=============================

Wide data has a column for each variable. For example, this is
wide-format data:

```{r, include=FALSE, echo=FALSE}
library(reshape2)
names(airquality) <- tolower(names(airquality))
aqm <- melt(airquality, id=c("month", "day"), na.rm=TRUE)
d <- dcast(aqm, month ~ variable, mean, margins = c("month", "variable"))
d <- d[1:4, -c(1, 3, 6)]
```

```{r, echo=FALSE}
head(d)
```

And this is long-format data:

```{r, echo=FALSE}
melt(d)
```

Long-format data has a column for possible variable types and a column
for the values of those variables. Long-format data isn’t necessarily
only two columns. For example, we might have ozone measurements for each
day of the year. In that case, we could have another column for day. In
other words, there are different levels of “longness”. The ultimate
shape you want to get your data into will depend on what you are doing
with it.

It turns out that you need wide-format data for some types of data analysis
and long-format data for others. In reality, you need long-format data much
more commonly than wide-format data. For example, `ggplot2` requires
long-format data 
(technically [tidy data](http://vita.had.co.nz/papers/tidy-data.html)),
`plyr` requires long-format data, and most modelling functions (such as
`lm()`, `glm()`, and `gam()`) require long-format data. But people often find
it easier to record their data in wide format.

The reshape2 package 
======================

`reshape2` is based around two key functions: `melt` and `cast`:

`melt` takes wide-format data and melts it into long-format data.

`cast` takes long-format data and casts it into wide-format data.

Think of working with metal: if you melt metal, it drips and becomes
long. If you cast it into a mould, it becomes wide.

Wide- to long-format data: the melt function 
==============================================

For this example we’ll work with the `airquality` dataset that is built
into `R`. First we’ll change the column names to lower case to make them
easier to work with. Then we’ll look at the data:

```{r}
names(airquality) <- tolower(names(airquality))
head(airquality)
```

What happens if we run the function `melt` with all the default argument
values?

```{r}
aql <- melt(airquality) # [a]ir [q]uality [l]ong format
head(aql)
```

```{r}
tail(aql)
```

By default, `melt` has assumed that all columns with numeric values are
variables with values. Often this is what you want. Maybe here we want
to know the values of `ozone`, `solar.r`, `wind`, and `temp` for each
`month` and `day`. We can do that with `melt` by telling it that we want
`month` and `day` to be “ID variables”. ID variables are the variables
that identify individual rows of data.

```{r}
aql <- melt(airquality, id.vars = c("month", "day"))
head(aql)
```

What if we wanted to control the column names in our long-format data?
`melt` lets us set those too all in one step:

```{r}
aql <- melt(airquality, id.vars = c("month", "day"),
  variable.name = "climate_variable", 
  value.name = "climate_value")
head(aql)
```

Long- to wide-format data: the cast functions 
===============================================

Whereas going from wide- to long-format data is pretty straightforward,
going from long- to wide-format data can take a bit more thought. It
usually involves some head scratching and some trial and error for all
but the simplest cases. Let’s go through some examples.

In `reshape2` there are multiple `cast` functions. Since you will most
commonly work with `data.frame` objects, we’ll explore the
`dcast` function. (There is also `acast` to return a vector, matrix, or
array.)

Let’s take the long-format `airquality` data and `cast` it into some
different wide formats. To start with, we’ll recover the same format we
started with and compare the two.

`dcast` uses a formula to describe the shape of the data. The arguments
on the left refer to the ID variables and the arguments on the right
refer to the measured variables. Coming up with the right formula can
take some trial and error at first. So, if you’re stuck don’t feel bad
about just experimenting with formulas. There are usually only so many
ways you can write the formula.

Here, we need to tell `dcast` that `month` and `day` are the ID
variables (we want a column for each) and that `variable` describes the
measured variables. Since there is only one remaining column,
`dcast` will figure out that it contains the values themselves. We could
explicitly declare this with `value.var`. (And in some cases it will be
necessary to do so.)

```{r}
aql <- melt(airquality, id.vars = c("month", "day"))
aqw <- dcast(aql, month + day ~ variable)
head(aqw)
```

```{r}
head(airquality) # original data
```

So, besides re-arranging the columns, we’ve recovered our original data.

If it isn’t clear to you what just happened there, then have a look at
this illustration:

![](/images/dcast-illustration.png)

Figure 1: An illustration of the `dcast` function. The blue shading indicates
ID variables that we want to represent individual rows. The red shading
represents variable names that we want to swing into column names. The grey
shading represents the data values that we want to fill in the cells with.

One confusing “mistake” you might make is casting a dataset in which
there is more than one value per data cell. For example, this time we
won’t include `day` as an ID variable:

```{r}
dcast(aql, month ~ variable)
```

When you run this in `R`, you’ll notice the warning message:

    # Aggregation function missing: defaulting to length

And if you look at the output, the cells are filled with the number of
data rows for each month-climate combination. The numbers we’re seeing
are the number of days recorded in each month. When you `cast` your data
and there are multiple values per cell, you also need to tell
`dcast` how to aggregate the data. For example, maybe you want to take
the `mean`, or the `median`, or the `sum`. Let’s try the last example,
but this time we’ll take the mean of the climate values. We’ll also pass
the option `na.rm = TRUE` through the `...` argument to remove `NA`
values. (The `...` let’s you pass on additional arguments to your
`fun.aggregate` function, here `mean`.)

```{r}
dcast(aql, month ~ variable, fun.aggregate = mean, 
  na.rm = TRUE)
```

Unlike `melt`, there are some other fancy things you can do with
`dcast` that I’m not covering here. It’s worth reading the help file
`?dcast`. For example, you can compute summaries for rows and columns,
subset the columns, and fill in missing cells in one call to `dcast`.

Additional help 
===============

Read the package help:
`help(package = "reshape2")`

See the `reshape2` website:
<http://had.co.nz/reshape/>

And read the paper on `reshape`:
Wickham, H. (2007). Reshaping data with the `reshape` package.
21(12):1–20.
<http://www.jstatsoft.org/v21/i12>  
(But note that the paper is written for the `reshape` package not the
`reshape2` package.)