### 2013-10-22 _knitr-posts/2013-10-21-panel-letters.Rmd
Add panel letters post draft

---
layout: post
title: Labelling Plot Panels
subtitle: Strategies for placing letters in figure panels.
---

Cole Monnahan writes:

> I've got a multipanel plot and simply want to add "(a)", "(b)", etc. to it.
> I've always cheated and used legend:
>
> ```
> add.label <- function(label, ...) legend("topleft", legend=" ", title=label,
> bty='n', ...)
> ```
>
> which works great for a topleft position. But for bottomright it isn't
> right. I can adapt it to:
>
> ```
> add.label2 <- function(label, position="topleft", ...) legend(x=position,
> legend=label, bty='n', ...)
> ```
>
> and that works great, but then doesn't work for topleft!
>
> There has got to be a non-hack way to do this simple (and common) task.

I'd never thought of using `legend()` for that purpose, but that's a neat
hack.

I tend to use `mtext()` for quick cases where the labels are all the same
width. For example if they're single letters, then you can align the labels
like this:

```{r panel_letters1, fig.width=6, fig.height=3}
par(mfrow = c(1, 2), mar = c(2, 2.5, 0, 0))
for(i in 1:2) {
  plot(1)
  mtext(paste0("(", letters[i], ")"), side = 3, adj = 0.05, 
    line = -1.3)
}
```

But the labels won't align if they are different widths:

```{r panel_letters2, fig.width=6, fig.height=3}
labels <- c("Short label", "A longer unaligned panel label")
par(mfrow = c(1, 2), mar = c(2, 2.5, 0, 0))
for(i in 1:2) {
  plot(1)
  mtext(labels[i], side = 3, adj = 0.05, line = -1.3)
}
```

And we can use `text()` combined with `par("usr")` to align these labels.
`par("usr")` gives us the coordinates of the plotting region in the order `x1,
x2, y1, y2`. Therefore, we can do the following to place the labels 2% over
from the left and 7% down from the top:

```{r panel_letters3, fig.width=6, fig.height=3}
labels <- c("Short label", "A longer panel label aligned")
par(mfrow = c(1, 2), mar = c(2, 2.5, 0, 0))
for(i in 1:2) {
  plot(1)
  u <- par("usr")
  x <- u[1] + 0.02 * (u[2] - u[1])
  y <- u[4] - 0.07 * (u[4] - u[3])
  text(x, y, labels[i], pos = 4)
}
```

And we could turn that into a little function:

```{r}
#' @param xfrac The fraction over from the left side.
#' @param yfrac The fraction down from the top.
#' @param label The text to label with.
#' @param ... Anything extra to pass to text(), e.g. cex, col.
add_label <- function(xfrac, yfrac, label, ...) {
  u <- par("usr")
  x <- u[1] + xfrac * (u[2] - u[1])
  y <- u[4] - yfrac * (u[4] - u[3])
  text(x, y, label, pos = 4, ...)
}
```

And then use it like this:

```{r panel_letters4, fig.width=6, fig.height=3}
labels <- c("Short", "A longer panel label")
par(mfrow = c(1, 2), mar = c(2, 2.5, 0, 0))
for(i in 1:2) {
  plot(1)
  add_label(0.02, 0.07, labels[i])
}
```