### 2014-10-13 _knitr-posts/2014-10-11-tmb.Rmd
Add draft TMB intro post

---
layout: post
title: The TMB R package
subtitle: Life beyond ADMB
published: false
---

I'd heard whispers of this new R package [TMB](https://github.com/kaskr/adcomp) and finally spent some time playing with it in the last few days. It's incredible — fast like ADMB (faster?), has built-in parallel functionality, has other advanced [statistical niceties](https://github.com/kaskr/adcomp/wiki), and it's built around R. What more could you ask for? (Well better documenation and the option to switch on Bayesian MCMC like you can in ADMB would be nice.)

The basic idea is to write a C++ template file and then compile and run it with some R functions. If you've used ADMB then the format will look similar. If you haven't, it's not too hard to pick up. It helps to [know a bit about C++](http://adv-r.had.co.nz/Rcpp.html), but for the most part it shouldn't be that hard for anyone familiar with R to pick it up.

The steps to running a TMB model are:

1. Write a C++ model template (a `.cpp` file).

2. Compile the model with `TMB::compile()`.

3. Load the model object with `dyn.load(TMB::dynlib())`

4. Make an object with data and parameters to pass to the TMB model with 
   `TMB::MakeADFun()`. If there are random effects / latent variables / unobserved
   states / coefficients that are allowed to vary (choose your preferred 
   terminology) then set the `random` argument in `TMB::MakeADFun()` to specify
   the name(s) of the parameters to integrate out of the likelihood.

5. Minimize the objective function with `optim()`.

6. Extract the estimates and standard errors from the model fit with 
   `TMB::sdreport()` and `TMB::summary.sdreport()`.

You can install the TMB package from GitHub:

```{r, eval=FALSE}
# install.packages("devtools") # if needed
devtools::install_github("kaskr/adcomp", subdir = "TMB")
```

The documentation is still very sparse, but you can find some instructions on [the R package GitHub wiki](https://github.com/kaskr/adcomp/wiki) and documentation of the underlying C++ functions and classes [here](http://folk.uib.no/hsk021/tmbdoc/).

## A linear regression example

Let's start with a really simple linear regression model.

We'll start by writing the TMB template file. Here, I've written it as inline R code that I then `write()` to a `.cpp` file. I have only done this so the code shows up here.

```{r tmb-regression}
tmb_model <- "
// linear regression
#include <TMB.hpp>

template<class Type>
Type objective_function<Type>::operator() () {
// data:
DATA_VECTOR(x);
DATA_VECTOR(y);

// parameters:
PARAMETER(a); // intercept
PARAMETER(b); // slope
PARAMETER(log_sigma); // log(residual SD)
// we fit sigma on a log scale to keep it > 0

// procedures: (transformed parameters)
Type sigma = exp(log_sigma);

int n = y.size(); // get number of data points to loop over

Type nll = 0.0; // initialize negative log likelihood

for(int i = 0; i < n; i++){ // C++ starts loops at 0!
  // get negative log likelihood (last argument is log = TRUE)
  nll -= dnorm(y[i], a + b * x[i], sigma, true);
}

return nll;
}"
write(tmb_model, file = "regression.cpp")
```

We'll load the model template:

```{r load, message=FALSE, results='hide'}
library(TMB)
compile("regression.cpp")
dyn.load(dynlib("regression"))
```

And simulate some data:

```{r tmb1, fig.height=5}
set.seed(1)
x <- runif(50, 1, 40)
y <- rnorm(50, mean = 2.5 + 0.3 * x, sd = exp(0.6))
plot(x, y)
```

Then we'll fit the model:

```{r, message=FALSE, results='hide', warning=FALSE}
obj <- MakeADFun(
  data = list(x = x, y = y), 
  parameters = list(a = 0, b = 0, log_sigma = 0),
  DLL = "regression")
opt <- do.call("optim", obj)
rep <- sdreport(obj)
```

We can check the estimates against the true values above (`a = 2.5, b = 0.3, log_sigma = 0.6`):

```{r}
rep
```

Which is exactly the same as we'd get from `stats::lm()`:

```{r lm}
arm::display(lm(y ~ x))
```

## Gompertz state space model

Now let's try something a bit more exciting — a state space Gompertz population dynamics model. We'll have TMB integrate out the unobserved "true" state that is masked by observation error. You can represent the Gompertz model as:

\\[\begin{aligned}
y_t &= \ln N\_t\\\      
U\_t &amp;= a + b U\_{t-1} + \epsilon\_t\\\
\epsilon_t &amp;\sim \mathrm{Normal}(0, \sigma^2\_\mathrm{proc})\\\
y\_t &amp;\sim \mathrm{Normal}(U\_t, \sigma^2\_\mathrm{obs}),
\end{aligned}\\]

where \\(y\_t\\) is the observed \\(\ln\\) abundance (\\(N\\)) at time \\(t\\). The model is density independent if \\(b = 1\\), maximally density dependent if \\(b = 0\\), and inversely density dependence if \\(b < 0\\). The parameter \\(a\\) represents the expected ln abundance at the next time step when \\(y\_t = 0\\). The process noise is modelled as a normal distribution with mean of \\(0\\), and a standard deviation of \\(\sigma\_\mathrm{proc}\\). \\(U\\) represents the unobserved state vector, and \\(\sigma\_\mathrm{obs}\\) represents the standard deviation of observation error (on a log scale).

The TMB model for this state-space Gompertz model is:
```{r tmb-model}
tmb_model <- "
// State-space Gompertz model
#include <TMB.hpp>

template<class Type>
Type objective_function<Type>::operator() () {
// data:
DATA_VECTOR(y);

// parameters:
PARAMETER(a); // population growth rate parameter
PARAMETER(b); // density dependence parameter
PARAMETER(log_sigma_proc); // log(process SD)
PARAMETER(log_sigma_obs); // log(observation SD)
PARAMETER_VECTOR(u); // unobserved state vector

// procedures: (transformed parameters)
Type sigma_proc = exp(log_sigma_proc);
Type sigma_obs = exp(log_sigma_obs);

// reports on transformed parameters:
ADREPORT(sigma_proc)
ADREPORT(sigma_obs)

int n = y.size(); // get time series length

Type nll = 0.0; // initialize negative log likelihood

// process model:
for(int i = 1; i < n; i++){
  Type m = a + b * u[i - 1]; // Gompertz
  nll -= dnorm(u[i], m, sigma_proc, true);
}

// observation model:
for(int i = 0; i < n; i++){
  nll -= dnorm(y[i], u[i], sigma_obs, true);
}

return nll;
}"
write(tmb_model, file = "gompertztmb.cpp")
```

We'll compile the model:

```{r compile, message=FALSE, results='hide'}
compile("gompertztmb.cpp")
dyn.load(dynlib("gompertztmb"))
```

And write a function that simulates some data, fits the model, and plots the estimates:

```{r fitfunction}
sim_gomp <- function(N = 100, seed = 123, sigma_obs = 0.2,
  sigma_proc = 0.2, a = 1.4, b = 0.7, y1 = 4) {

  set.seed(seed)
  N <- N
  ytrue <- numeric(length = N)
  ytrue[1] <- y1
  log_sigma_proc <- log(sigma_proc)
  proc_error <- rnorm(N, mean = 0, sd = sigma_proc)
  log_sigma_obs <- log(sigma_obs)
  for(i in 2:N) {
    ytrue[i] <- a + b * ytrue[i-1] + proc_error[i-1]
    }
  x <- seq_len(N)
  y <- rnorm(N, mean = ytrue, sd = sigma_obs)

  data <- list(y = y)
  parameters <- list(a = 1, b = 0.5, log_sigma_proc = -1,
    log_sigma_obs = -1, u = rep(mean(y), N))
  obj <- MakeADFun(data, parameters, random = "u", DLL = "gompertztmb")
  obj$hessian <- FALSE
  opt <- do.call("optim", obj)
  rep <- sdreport(obj)

  # extract estimated process:
  u <- summary(rep, "random")[, "Estimate"]
  u_se <- summary(rep, "random")[, "Std. Error"]
  # extract fixed effects:
  fixed <- summary(rep, "fixed")

  # We'll write a little helper function to plot fixed effect 
  # estimates:
  plot_fixed <- function(par, y, mat, trans = I) {
    points(trans(eval(parse(text = par))), y, pch = 21,
      bg = "#00000030", col = "#00000080")
    f <- mat[par, "Estimate"]
    se <- mat[par, "Std. Error"]
    points(trans(f), y, pch = 19, col = "red")
    segments(trans(f - 2 * se), y, trans(f + 2 * se), y, pch = 19,
      col = "red")
    }

  par(mfrow = c(2, 1), mar = c(4, 6, 0, 0), oma = c(.5, .5, .5, .5),
    mgp = c(2, 0.5, 0), cex = 1, tck = -0.02)
  plot(1, 1, xlim = c(0, 2), ylim = c(1, 4), type = "n",
    xlab = "Coefficient value", ylab = "", yaxt = "n")
  axis(2, at = 1:4, labels = c("a", "b", "sigma_obs", "sigma_proc"),
    las = 1)

  plot_fixed("a", 1, fixed)
  plot_fixed("b", 2, fixed)
  plot_fixed("log_sigma_obs", 3, fixed, trans = function(z) exp(z))
  plot_fixed("log_sigma_proc", 4, fixed, trans = function(z) exp(z))

  plot(x, y, col = "#00000080", pch = 21, bg = "#00000030", las = 1,
    ylab = "log(abundance)", xlab = "time")
  lines(x, ytrue, lty = 2, lwd = 2)
  lines(x, u, col = "red", lty = 1, lwd = 1.5)
  polygon(c(x, rev(x)), c(u - 2 * u_se, rev(u + 2 * u_se)), 
    border = NA, col = "#FF000050")
  legend("bottomright", legend = c("Observed", "True", "Estimated"),
    pch = c(21, NA, NA), lty = c(NA, 2, 1), lwd = c(NA, 2, 1.5),
    col = c("#00000080", "black", "red"), 
    pt.bg = c("#00000030", NA, NA))
  }
```

One good test that you've coded a model correctly is to throw a boat-load of data at it and make sure it comes back with unbiased parameter estimates. In the following plots, the first panel shows the coefficient estimates ("fixed effects") +/- two standard errors. The red dots are our estimates and the shaded grey dots are the true values. In the lower panels, the red lines are our estimated states, the black line (hidden behind the red in the first plot) is the true underlying state, and the grey dots are our observations.

```{r tmb-test1, results='hide', message=FALSE, warning=FALSE, fig.height=8}
sim_gomp(N = 5000, sigma_obs = 0.1)
```

(We just modelled a 5000 time step state space model in a couple of seconds!)

Now we can try a more realistic 50 data points:

```{r tmb-short, results='hide', message=FALSE, warning=FALSE, fig.height=8}
sim_gomp(N = 50)
```

And with a different seed value but more data:

```{r tmb-seed2, results='hide', message=FALSE, warning=FALSE, fig.height=8}
sim_gomp(N = 100, seed = 999)
```