### 2015-03-12 _knitr-posts/2015-02-02-stan-matt-trick.Rmd
Add 2 draft posts

---
layout: post
title: The Matt Trick in Stan
subtitle: What do they mean?
published: false
---

If you read about Stan for a while you'll invariably hear about the 'Matt Trick'. In fact, most of the example models coded in the GitHub wiki are coded with the Matt Trick. The Matt Trick is a way of writing a multilevel/hierarchical/mixed-effects model in Stan to create less-correlated, better-mixing sample chains. Thing is, I can't find one good explanation anywhere what the trick is, and every time I do it I have to cobble together bits and pieces of information. Then I quickly forget what I was doing. This post is mostly to document the 'trick' for myself as a reference.

We'll use a 'simple' multilevel model with one predictor and varying intercepts. We'll start by simulating some data, then we'll fit the model normally and fit the model with the 'trick'. Then we'll look at run times and mixing properties of the chains.

```{r}
set.seed(123)
N <- 50 # number of observations per group
Nj <- 16 # number of groups
sigma_j <- 2.0 # standard deviation of intercepts across groups
d <- data.frame(j = 1:Nj, jname = letters[1:Nj], b0_j = rnorm(Nj, 0, sigma_j))
sigma <- 1.1 # residual standard deviation
b0 <- 0.3 # intercept
b1 <- 0.5 # slope
x <- lapply(seq_len(Nj), function(x) rnorm(n = N)) # predictor
dat <- list()
for(j in d$j) {
  dat[[j]] <- data.frame(j = j, jname = d$jname[j], x = x[[j]],
    y = rnorm(N, mean = b0 + x[[j]] * b1 + d$b0_j[j], sd = sigma))
}
dat <- as.data.frame(do.call("rbind", dat))
library("ggplot2")
ggplot(dat, aes(x, y)) + geom_point() + facet_wrap(~jname)
```

The model coded normally:

```{r}
model <- "
data {
  int<lower=1> N;        // rows of data
  int<lower=1> Nj;      // number of groups
  int<lower=1,upper=Nj> j_id[N];  // group ids
  vector[N] x;       // predictor
  real y[N]; // response
}
parameters {
  real<lower=0> sigma;  // residual standard deviation
  real b0;              // intercept
  real b1;              // slope
  real<lower=0> sigma_j; // group-level standard deviation
  vector[Nj] b0_j;
}
transformed parameters {
  vector[N] yhat;
  for (i in 1:N) {
    yhat[i] <- b0_j[j_id[i]] + b1 * x[i];
  }
}
model {
  b0_j ~ normal(b0, sigma_j); 
  // priors:
  sigma ~ cauchy(0, 2.5);
  sigma_j ~ cauchy(0, 2.5);
  b0 ~ normal(0, 2.5);
  b1 ~ normal(0, 2.5);
  y ~ normal(yhat, sigma);
}
"
sm <- rstan::stan_model(model_code = model)
```

The model with the Matt Trick:

```{r}
model <- "
data {
  int<lower=1> N;                 // rows of data
  int<lower=1> Nj;                // number of groups
  int<lower=1,upper=Nj> j_id[N];  // group ids
  vector[N] x;                    // predictor
  real y[N];                      // response
}
parameters {
  real<lower=0> sigma;            // residual standard deviation
  real b0;                        // intercept
  real b1;                        // slope
  real<lower=0> sigma_j;          // group-level standard deviation
  vector[Nj] e_b0_j;              // Matt Trick #1
}
transformed parameters {
  vector[N] yhat;
  vector[Nj] b0_j;                // Matt Trick #2
  b0_j <- b0 + sigma_j * e_b0_j;  // Matt Trick #3

  for (i in 1:N) {
    yhat[i] <- b0_j[j_id[i]] + b1 * x[i];
  }
}
model {
  e_b0_j ~ normal(0, 1);          // Matt Trick #4
  sigma ~ cauchy(0, 2.5);
  sigma_j ~ cauchy(0, 2.5);
  b0 ~ normal(0, 2.5);
  b1 ~ normal(0, 2.5);
  y ~ normal(yhat, sigma);
}
"
sm_matt <- rstan::stan_model(model_code = model)
```

Now let's fit the models:

```{r}
model_data <- list(N = nrow(dat), y = dat$y, x = dat$x, Nj = max(dat$j), 
  j_id = dat$j)
pars <- c("b0", "b1", "sigma", "sigma_j")

st <- system.time({
  m <- sampling(sm, data = model_data, iter = 400, chains = 3, 
    pars = pars, seed = 999)
})
st_matt <- system.time({
  m_matt <- sampling(sm_matt, data = model_data, iter = 400, chains = 3, 
    pars = pars, seed = 999)
})
```