### 2017-02-17 _knitr-posts/2017-02-17-negative-binomial.Rmd
Add draft negbin post

---
layout: post
title: 
subtitle: 
published: false
---

> A group of us [...] are in a "Bayesian club" and trying to use our collective
> intelligence to learn more about Bayesian modelling. [...] Today we decided to
> overview probability distributions [...] identifying the different main
> distribution types and how they are parameterized. It all went pretty good and
> we were all pretty clear on things... until we got to the negative binomial
> distribution. The bit of research/reading and googling that we had done
> suggested this is a notoriously nebulous distribution in that there are
> multiple ways it can be parameterized? [...]

There certainly are a lot of ways to parameterize the negative binomial. I've
seen claims of up to 7 parameterizations. Not only are there different ways to
parameterize it but there are at least 2 different versions of the negative
binomial distribution itself. One where the variance scales quadratically with
the mean and another where it scales linearly. By far the most common in
ecology is the version where the variance scales quadratically with the mean.
This is often referred to as the NB2 in negative binomial. I think that name
was coined in this book:
https://books.google.com/books/about/Negative_Binomial_Regression.html?id=DDxEGQuqkJoC

The other version where the variance and mean scale linearly is sometimes
called NB1. In general, I wouldn't worry too much about this besides being able
to describe the version of the distribution that you have used. I'd focus on
the NB2.

As with any distribution, you need to read the manual or help files if you are
using a probability density function written by someone else. A common
parameterization in ecology is one that is parameterized in terms of a mean and
a dispersion parameter, sometimes called phi, theta, or size.

In that case, you have NB2(mu, phi), and the mean and the variance at a given
mean value are:
mean = mu
variance = mu + (mu^2)/phi

So as phi becomes large the variance approaches the mean and the distribution
approaches the Poisson distribution.

This is how MASS::rnegbin is parameterized. Although they call the dispersion
parameter theta.

stats::dnbinom() can also be parameterized this way. Although they call the
dispersion parameter size.

This is also how the negative binomial is parameterized in glmmTMB and
glmmADMB.

And this is one of the ways it is parameterized in Stan:
neg_binomial_2(mu, phi)

Note that sometimes the distribution is parameterized as:
variance = mu + (mu^2) * phi

In which case, phi has simply been inverted so that as phi become small, the
variance approaches the mean and the distribution approaches the Poisson
distribution.

It's been a while since I worked in JAGS/BUGS, but after a quick look it seems
like JAGS is parameterized with 2 parameters we will call p and r. Then:
dnegbin(p, r)
mean = r(1-p)/p
variance = r(1-p)/p^2

Simple, right? :)  This matches the stats::dnbinom distribution if you use the
size and prob arguments... except that the order of the 2 parameters is
reversed. They wouldn't want to make it too easy after all. Details here:
https://journal.r-project.org/archive/2013-1/lebauer-dietze-bolker.pdf

Now if you want to write a GLM or GLMM yourself, you'll be working with a log
link and then exponentiating before passing the values to the density function.

I think, based on a bit of Googling, you could fit a model in JAGS with
something like this:

mu[i] <- b0 + b1 * x1[i] ...
lambda[i] <- exp(mu[i])
p[i] <- r/(r+lambda[i])
y[i] ~ dnegbin(p[i],r)

Then,
mean = lambda
variance = lambda + (lambda^2)/r

Stan has a built in parameterization on the log scale so that you can avoid the
exponentiating. This is a bit faster, cleaner, and less likely to run into
numerical issues with big or small numbers:

mu = b0 + b1 * x1...
y ~ neg_binomial_2_log(mu, phi)

If you're interested in fitting negative binomial GLMs and GLMMs in Stan,
I wrote these examples a couple years ago:
https://github.com/seananderson/negbin-stan/blob/master/negbin-glm.md

(You can also now fit these easily with the rstanarm package.)

As with any model you write yourself, it's important to check your model by
either comparing it with some other version that you trust from a different
package or software or by simulating some data and recovering the true values
with your model (as I was doing in the above link). This is especially so if
you're unsure about the the parameterization of a distribution.